Synthesizing Symmetric Lenses

### 2019-02-27 papers/SymmetricLenses/main.tex
fix issues

\lstset{ language=Caml, basicstyle=\upshape\sffamily,
keywordstyle=\upshape\sffamily\color{dkpurple}, keepspaces=true,
framexleftmargin=1ex, framexrightmargin=1ex, showstringspaces=true,
commentstyle=\itshape\rmfamily,
emph={rep,iterate,synth,collapse,perm,squash,normalize,using,ins,del,lens,let,get,put,rquot,lquot,id,swap,concat,or,disconnect,merge_left,merge_right,const},
emphstyle=\upshape\sffamily\color{dkpurple}, 
columns=fullflexible,
mathescape, 
xleftmargin=1.5em,
% BCP: I find this distracting:
stringstyle=\sffamily\color{dkblue},
aboveskip=.3em,
belowskip=.3em,
}
\makeatletter
     \let\lst@oldvisiblespace\lst@visiblespace
     \def\lst@visiblespace{\,\lst@oldvisiblespace\,}
\makeatother

@@

  \setbox0=\hbox{%
    \begin{minipage}{3.5in}
\begin{lstlisting}
let company = (name . ("Co." | "Inc." | "Ltd.")) | "UNK" in
let emp_ins = name . " " . name "," company in
let header = "FirstLast,Company" in
let emp_insurance = header . ("\n" . emp_ins)$^*$
\end{lstlisting}
    \end{minipage}
  }
  \savestack{\listingD}{\box0}

@@

Without unambiguity constraints, the method above for calculating entropy does
not work, and it is not easy to see how to generalize it.  For example,
consider the SRE $\PRegexOr{\lstinline{"a"}}{\lstinline{"a"}}{.5}$. Using
the above computation, this SRE has an entropy of $1$, but the entropy
should be $0$. These difficulties become particularly complex when the
languages are ambiguous in subtle ways and the algorithm must be correct
when only some of the two languages overlap. Similar issues occur when
trying to find the entropy when $\emptyset$ is a subterm (what should the
entropy of $\PRegexStar{\emptyset}{.5}$ be?), so we do not define entropy
on~$\emptyset$.

@@

\section{Simple Symmetric String Lenses}
\label{sec:ssl}
We now give a formal description of all the combinators in our symmetric lens
language. While our algorithms perform synthesis on stochastic regular
expressions, lenses themselves are typed between regular expressions. We can
recover a regular expression from a stochastic regular expression by deleting
probability annotations. In a slight abuse of notation (as we are also using
these symbols as nonterminals), if $\Regex$ is a
stochastic regular expression, $\BRegex$ is the regular expression obtained by
removing probability annotations. The typing judgment $\Lens \OfType \BRegex
\Leftrightarrow \BRegexAlt$ means that $\Lens$ is a well typed lens between
the language of \BRegex and the language of \BRegexAlt.

@@

We have already described a number of these combinators in \S\ref{sec:overview}.
Briefly (full definitions are below), the compose lens $\Lens_1 ; \Lens_2$
composes two lenses sequentially, it 
transforms from one data format to the other by first transforming to an
intermediate format, shared as the target of $\Lens_1$ and the source of
$\Lens_2$. The \MergeR lens combines two sublenses which share the same target.
Which lens gets used depends on what data is in the left-hand format, if it is
in the domain of $\Lens_1$ then $\Lens_1$ gets used, and similarly for
$\Lens_2$.  The \MergeL lens is symmetric to \MergeR, the two sublenses must
share the same source. Finally, $\InvertOf{\Lens}$ inverts a lens, \CreateR
becomes \CreateL and \PutR becomes \PutL, and vice-versa.

We now give typing rules and semantics for (some of) these combinators. We do not
include the semantics for combinators equivalent to those presented in prior
work~\cite{boomerang}. Furthermore, we present
just \CreateR and \PutR in cases where \CreateL and \PutL are symmetric;
full details can be found in an appendix\ifappendices \S\ref{sec:appendixlenses}  \else{} (in the supplemental materials)\fi. For
simplicity of presentation, we use unambiguous regular expressions
everywhere, even though the identity and disconnect lenses can actually be defined over ambiguous
regular expressions; we believe relaxing
this restriction is unproblematic.

@@

@@

\subsection{Lens Costs}
\label{subsec:lenscosts}
Our cost metric is based on the expected amount of information required to
recover a string in one data format from the other. We use the function
$\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ to calculate bounds of the
expected amount of information required to recover a string in $\RegexAlt$ from
a string in $\Regex$, synchronized by $\Lens$. Similarly, we use the function
$\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ to calculate bounds of the
expected amount of information required to recover a string in $\Regex$ from a
string in $\RegexAlt$, synchronized by $\Lens$. We write $a(b,c)$ to
mean $(a*b,a*c)$, and $(a,b)+(c,d)$ to mean $(a+c,b+d)$.
\begin{center}
  \begin{tabular}{rcl}
    $\REntropyOf{\RegexAlt \Given \IdentityLensOf{\RegexAlt}, \RegexAlt}$
    & =
    & [0,0]\\

@@

Note that our definition of $\Entropy^\rightarrow$ contains no case for sequential
composition $\Lens_1 ; \Lens_2$ and our theorem excludes lenses that contain
such compositions.  Defining the entropy of lenses involving
composition is challenging because $\Lens_1$ might, for instance, add some
information that is subsequently projected away in $\Lens_2$. Such operations
can cancel, leaving a zero-entropy bijection composed from two non-zero entropy
transformations. However, detecting such cancellations directly is complicated and
this property is difficult to determine merely from syntax. Fortunately, we are able to
sidestep such considerations by synthesizing \emph{DNF lenses}---simple symmetric
lenses that inhabit a disjunctive normal form.

@@

% begin synthesis
\section{Synthesis}
\label{sec:synthesis}
Algorithm~\ref{alg:synth-sym-lens} presents our synthesis algorithm at a high
level of abstraction. The input regular expressions are first converted into
stochastic regular expressions with \ToStochastic.  
%Once that conversion has
%taken place, the $\RXSearch$ algorithm executes. This algorithm itself is
%defined via five functions: (1) $\PCF{\PQ.Create}$, (2) $\PCF{\PQ.Pop}$, (3)
%$\PCF{Expand}$, (4) $\PCF{\PQ.Push}$, and (5) $\PCF{\Continue}$. First,
This pair of SREs is used to initialize a priority queue ($pq$). The priority of
a SRE pair is the number of rewrites needed to derive the pair from the
originals. Next, $\PCF{SynthSymLens}$ enters a loop that searches for likely
lenses. The loop terminates when the algorithm believes it is unlikely to find a
better lens than the best one it has found so far (a termination condition
defined by $\PCF{\Continue}$). Within each iteration of the loop, it:
\begin{itemize}
\item pops the next class ($S$, $T$) of lenses to
search off of the priority queue ($\PCF{\PQ.Pop}$),
\item executes $\GreedySynth$ to find a best lens in that class if
one exists ($\LC$),
using the examples $exs$ to filter out potential lenses that do not satisfy
the specification,
\item replaces $best$ with $\LC$, if $\LC$ has lower cost according
to our information-theoretic metric, and
\item adds the SREs derived from rewriting $S$ and $T$ ($\Expand(S,T)$)
  to the priority queue.
\end{itemize}
When the loop terminates, the search returns the globally best lens found
($best$).  Each subroutine of this algorithm will be explained in
further depth in the following subsections.

\emph{Stochastic DNF regular expressions} are critical to the success of our
algorithm. Stochastic DNF regular expressions (\SDNFREabbrev{}s) are a
pseudo-normal form for stochastic regular expressions. Of all the star-semiring
equations required to relate two \SDNFREabbrevs, only those relating to Kleene
star (\EG, \UnrollstarLeftRule{} and \UnrollstarLeftRule{}) are needed to relate
two \SDNFREabbrevs; other axioms, like associativity, commutativity or
distributivity, are unneeded. This important property makes it possible to
construct an efficient algorithm ($\GreedySynth$), directed by the syntax of two
\SDNFREabbrevs, to synthesize any lenses in the same ``class'' (\IE, any lenses
that have the same type when the Kleene star axioms are unused). This property,
and the resulting efficiency of $\GreedySynth$ is key to the effectiveness of
our core algorithms.

@@

The procedure \PCF{\Continue} returns false based on the how long the search has
been going, and how hard it expects the next class of problems to be. In
particular, if $\PCF{PQ.Peek}(pq) = (\Regex,\RegexAlt)$, that RE pair is at
distance $d$, the number of pairs in $pq$ at distance $d$ is $n$, and the
current best lens has cost $c$, then
  $\Call{Continue}{pq}$ checks whether $c > d + \Log_2(n)$.

When $\Call{Continue}{pq}$ returns $\False$, the algorithm stops proposing
regular expression pairs, and instead returns to the user the best lens found
thus far. If the algorithm finds a bijective lens, which
has zero cost, it will immediately return.
%
\subsection{Stochastic DNF Regular Expressions}
\label{subsec:sdnfre}
Just as SREs are REs with probability annotations,
\SDNFREabbrevs are \DNFREabbrevs with probability annotations.
As with SREs and regular
expressions, if $\DNFRegex$ is a stochastic DNF regular expression,
$\overline{\DNFRegex}$ is the DNF regular expression generated by erasing
probability annotations.

@@

The algorithm for converting a stochastic regular expressions into its
corresponding stochastic DNF regular expressions, written
$\ToDNFRegexOf{\Regex}$, is defined below.
\[
  \begin{array}{rcl@{\hspace*{3em}}rcl}
    \ToDNFRegexOf{\String} 
    & = 
    & \DNFOf{(\SequenceOf{\String},1)}

@@

  \end{array}
\]
This conversion relies on operators defined in
Figure~\ref{fig:dnf-regex-functions}. We have proved that this conversion
respects languages and probability distributions.

@@

\paragraph*{\ToStochastic} With stochastic DNF regular expressions and
\ToDNFRegex defined, it is easier to explain \ToStochastic, the function that converts regular
expressions into stochastic regular expressions. If users
provided stochastic regular expressions themselves (perhaps inferred from
example data), this step could be skipped,
though our implementation does not (yet) permit users to provide such
stochastic regular expressions. If $\Regex$ is a stochastic DNF regular
expression generated by \ToStochastic, then $\ToDNFRegexOf{\Regex} =
\DNFOf{(\Sequence_1,\frac{1}{n}) \DNFSep \ldots \DNFSep
  (\Sequence_n,\frac{1}{n})}$ for some sequences $\Sequence_1\ldots\Sequence_n$,
and every stochastic atom generated by \ToStochastic is
$\PRegexStar{\DNFRegex}{.8}$. In particular, $\ToDNFRegex$ generates regular
expressions whose DNF form gives equal probability to all sequence subcomponents
of the DNF regular expression, and gives iterations a .8 chance to continue
iterating. In our experience generating random strings from regular expressions,
these probabilities provide good distributions of strings---stars are iterated
4 times on average, and no individual choice in a series disjunctions is chosen
disproportionately often. As we have found these distributions generate
reasonable strings, we believe they describe the shape of data well.

@@

@@

Intuitively, a symmetric DNF lens $sdl$ is a union of symmetric sequence lenses,
$ssql_1 \ldots ssql_p$, a symmetric DNF sequence lens is a concatenation of
symmetric atom lenses, $sal_1 \ldots sal_n$, and a symmetric atom lens $sal$ is
an iteration of a symmetric DNF lens. Note that the structure of symmetric DNF
lenses exactly mirrors that of SDNF REs.

Just as we analyzed the information content of ordinary regular expressions, we
can analyze the information content of DNF regular expressions. As before, we
use $\REntropyOf{\DNFRegexAlt \Given \SDNFLens, \DNFRegex}$ to calculate bounds
of the expected amount of information required to recover a string in
$\DNFRegexAlt$ from a string in $\DNFRegex$, synchronized by $\SDNFLens$. We use
the function $\LEntropyOf{\DNFRegex \Given \SDNFLens, \DNFRegexAlt}$ to
calculate bounds on the expected amount of information required to recover a
string in $\DNFRegex$ from a string in $\DNFRegexAlt$, synchronized by
$\SDNFLens$.

@@

\paragraph*{Compositional Synthesis} Compositional synthesis allows us to use
previously defined lenses, whether they are previously defined by users or by
synthesis. As the synthesizer processes a Boomerang file, it accumulates
lens definitions and their types.  It tackles synthesis problems one after another
and there may be many such problems in a given program.
During synthesis, if an existing lens has the right type and agrees with the examples.
\GreedySynth will use it.  Consequently, if a large synthesis problem cannot be solved all
at once, a user can first generate subproblems for parts of a data format, and the
larger problem can subsequently use solutions to those subproblems.  In our experience,
this is a very powerful tool that allows synthesis to tackle problems of arbitrary
complexity.

In this work, there are three competing constraints. The tool should (1) return
lenses quickly to users, (2) return the highest quality lenses, and (3)
not require users to write a large number of examples.
Our termination heuristic works well for most situations, but
there are cases where it fails.

In particular, our system works hard to find lenses that project \emph{as little
  information as possible}. If the desired lens requires projecting a large
amount of information, the termination condition will be satisfied ``too late.''
While \GreedySynth may have found what the user knows to be the desired lens,
the full algorithm will continue trying to find lenses that project less
information.

Conversely, sometimes the termination heuristic is overly permissive and algorithm cuts
short: if it 
had continued searching for lenses it would have found the desired one, but the
termination heuristic was satisfied ``too early.'' We wish for our algorithm to
provide feedback in a timely manner, but sometimes this prohibits finding the
desired lens.

To address these issues, we allow annotations on the regular expressions
dictating relevance. The first, $\SkipOf{\Regex}$, says the information
content of $\Regex$ is irrelevant to the desired lens, and can the data can be
safely projected. The second, $\SRequireOf{\Regex}$, says the information of
$\Regex$ is critical to the desired lens, and none of the information can be
lost when converting between the formats.

If synthesizing a lens seems to be taking a long time, it is typically spending
time checking it cannot find a lens that projects less information. By
``skipping'' portions of the SRE that \emph{should} be projected, the algorithm
will not continue searching for a better lens when it has already found the
correct one. In practice, we implement \Skip by setting the information content
of $\Skip(\Regex)$ to zero.
%
\begin{gather*}
  \EntropyOf{\Skip(\Regex)} = 0
\end{gather*}
%
If synthesizing a lens between \lstinline{emp_salaries} and
\lstinline{emp_insurance} from Figure~\ref{fig:minimized-representations} is
taking overly long, the synthesis procedure is uncomfortable disconnecting
\lstinline{company} and \lstinline{salary} information. Redefining
\lstinline{company} and \lstinline{salary} as $\SkipOf{\lstinline{company}}$ and
$\SkipOf{\lstinline{salary}}$ would set their information content to zero,
alleviating such issues.

If the algorithm returns a lens with too many disconnects and merges, one could
provide additional examples to constrain the lens more. However, for large
formats, writing such examples is time consuming and error-prone. Instead, if
one recognizes that information from $\Regex$ is incorrectly lost when converted
between formats, they can prevent that information from being lost by annotating
it $\SRequireOf{\Regex}$. One can guarantee a lens between \Regex and \RegexAlt
is bijective by annotating them as $\SRequireOf{\Regex}$ and
$\SRequireOf{\RegexAlt}$. In practice, we implement \SRequire by setting the
cost of any lens that loses ``required'' information to infinity.
%
\begin{center}
  \begin{tabular}{rcl}
    $\REntropyOf{\SRequireOf{\RegexAlt} \Given \Lens, \Regex}$
    & =
    & $\begin{cases}
      \infty & \text{if }\REntropyOf{\SRequireOf{\RegexAlt} \Given \Lens, \Regex}
      \neq 0\\
    0 & \text{otherwise}
  \end{cases}$\\

    $\REntropyOf{\RegexAlt \Given \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}, \Regex}$
    & =
    & $\begin{cases}
      \infty & \vspace*{-.3em}\text{if $\RegexAlt$ contains $\SRequireOf{\RegexAlt'}$ as a subexpression,}\\
      & \text{where $\EntropyOf{\RegexAlt'} \neq 0$}\\
    [\EntropyOf{\RegexAlt},\EntropyOf{\RegexAlt}] & \text{otherwise}
  \end{cases}$\\
  \end{tabular}
\end{center}

%
If synthesizing a lens between \lstinline{emp_salaries} and
\lstinline{emp_insurance} from Figure~\ref{fig:minimized-representations}
quickly returns a lens with too many disconnects and merges (perhaps
disconnecting instances of \lstinline{name} in \lstinline{emp_salary} and
\lstinline{emp_ins}), the termination heuristic is being overly permissive. By
replacing \lstinline{name} with $\SRequireOf{\lstinline{name}}$ in
\lstinline{emp_salary} and \lstinline{emp_ins}, name information can no longer
be lost, and future synthesis calls would continue until a more likely lens is
found.
% end synthesis

>>>>

\lstset{ language=Caml, basicstyle=\upshape\sffamily,
keywordstyle=\upshape\sffamily\color{dkpurple}, keepspaces=true,
framexleftmargin=1ex, framexrightmargin=1ex, showstringspaces=true,
commentstyle=\itshape\rmfamily,
emph={rep,iterate,synth,collapse,perm,squash,normalize,using,ins,del,lens,let,get,put,rquot,lquot,id,swap,concat,or,disconnect,merge_left,merge_right,const,skip,require},
emphstyle=\upshape\sffamily\color{dkpurple}, 
columns=fullflexible,
mathescape, 
xleftmargin=1.5em,
% BCP: I find this distracting:
stringstyle=\sffamily\color{dkblue},
aboveskip=.3em,
belowskip=.3em,
}
\makeatletter
     \let\lst@oldvisiblespace\lst@visiblespace
     \def\lst@visiblespace{\,\lst@oldvisiblespace\,}
\makeatother

@@

  \setbox0=\hbox{%
    \begin{minipage}{3.5in}
\begin{lstlisting}
let company = (co_name . ("Co." | "Inc." | "Ltd.")) | "UNK" in
let emp_ins = name . " " . name "," company in
let header = "FirstLast,Company" in
let emp_insurance = header . ("\n" . emp_ins)$^*$
\end{lstlisting}
    \end{minipage}
  }
  \savestack{\listingD}{\box0}

@@

To understand the difficulties caused by ambiguity, consider the SRE
$\PRegexOr{\lstinline{"a"}}{\lstinline{"a"}}{.5}$. The formula above defines the
entropy to be 1, but the true entropy is 0 (\IE, no bits are needed to know the
generated string will be \lstinline{"a"}). Similar issues occur when trying to
find the entropy when $\emptyset$ is a subterm (what should the entropy of
$\PRegexStar{\emptyset}{.5}$ be?), so we do not define entropy on~$\emptyset$.

@@

\section{Simple Symmetric String Lenses}
\label{sec:ssl}
Simple symmetric string lenses are bidirectional programs that convert string
data between formats. These lenses operate over regular expressions, not SREs.
If $\Regex$ is a stochastic regular expression, $\BRegex$ is the regular
expression obtained by removing probability annotations.

@@

We have already described a number of these combinators in \S\ref{sec:overview}.
Briefly (full definitions are below), the lens $\Lens_1 ; \Lens_2$
composes two lenses sequentially, it 
transforms from one data format to the other by first transforming to an
intermediate format, shared as the target of $\Lens_1$ and the source of
$\Lens_2$. The \MergeR lens combines two sublenses that operate on disjoint
source formats and share the same target.
The lens that gets used depends on what data is in the left-hand format, if the data is
in the domain of $\Lens_1$ then $\Lens_1$ gets used, and similarly for
$\Lens_2$.  The \MergeL lens is symmetric to \MergeR, the two sublenses must
share the same source. Finally, $\InvertOf{\Lens}$ inverts a lens, \CreateR
becomes \CreateL and \PutR becomes \PutL, and vice-versa.

We now give typing rules and semantics for (some of) these combinators. We write
$\Lens \OfType \BRegex \Leftrightarrow \BRegexAlt$ to indicate that $\Lens$ is a
well typed lens between the language of \BRegex and the language of \BRegexAlt.
We present just \CreateR and \PutR in cases where \CreateL and \PutL are
symmetric; full details can be found in an appendix\ifappendices
\S\ref{sec:appendixlenses} \else{} (in the supplemental materials)\fi. For
simplicity of presentation, we use unambiguous regular expressions everywhere,
even though the identity and disconnect lenses can actually be defined over
ambiguous regular expressions; relaxing this restriction is not problematic.

@@

Just as the identity lens ignores the second argument in puts, disconnect lenses
ignore the first argument in both puts and creates.  The data is unsynchronized in these
two formats, information from one format doesn't impact the other.

@@

\subsection{Lens Costs}
\label{subsec:lenscosts}
Our cost metric is based on the expected amount of information required to
recover a string in one data format from the other. We use the function
$\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ to calculate bounds on the
expected amount of information required to recover a string in $\RegexAlt$ from
a string in $\Regex$, synchronized by $\Lens$. Similarly, we use the function
$\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ to calculate bounds on the
expected amount of information required to recover a string in $\Regex$ from a
string in $\RegexAlt$, synchronized by $\Lens$. We write $a(b,c)$ to
mean $(a*b,a*c)$, and $(a,b)+(c,d)$ to mean $(a+c,b+d)$.
\begin{center}
  \begin{tabular}{rcl}
    $\REntropyOf{\RegexAlt \Given \IdentityLensOf{\RegexAlt}, \RegexAlt}$
    & =
    & [0,0]\\

@@

Note that our definition of $\Entropy^\rightarrow$ contains no case for sequential
composition $\Lens_1 ; \Lens_2$ and our theorem excludes lenses that contain
such compositions.  Defining the entropy of lenses involving
composition is challenging because $\Lens_1$ might, for instance, add some
information that is subsequently projected away in $\Lens_2$. Such operations
can cancel, leaving a zero-entropy bijection composed from two non-zero entropy
transformations. However, detecting such cancellations directly is complicated and
this property is difficult to determine merely from syntax. Fortunately, we are able to
sidestep such considerations by synthesizing \emph{DNF lenses}---simple symmetric
lenses that inhabit a disjunctive normal form that does not include composition.

@@

% begin synthesis
\section{Synthesis}
\label{sec:synthesis}
Algorithm~\ref{alg:synth-sym-lens} presents our synthesis algorithm at a high
level of abstraction. This algorithm searches for likely lenses in priority
order one ``class'' at a time using a \GreedySynth subroutine. Each class is the
set of lenses that can by typed by a pair of regular expressions, modulo a set
of simple axioms such as associativity, commutativity, and distributivity. The
\Expand subroutine generates new classes using the star-unrolling axioms
(\UnrollstarLeftRule and \UnrollstarRightRule).

To summarize, the input regular
expressions are first converted into stochastic regular expressions with
\ToStochastic. This pair of SREs is used to initialize a priority queue ($pq$).
The priority of a SRE pair is the number of rewrites needed to derive the pair
from the originals. Next, $\PCF{SynthSymLens}$ enters a loop that searches for
likely lenses. The loop terminates when the algorithm believes it is unlikely to
find a better lens than the best one it has found so far (a termination
condition defined by $\PCF{\Continue}$). Within each iteration of the loop, it:
\begin{itemize}
\item pops the next class ($S$, $T$) of lenses to
search off of the priority queue ($\PCF{\PQ.Pop}$),
\item executes $\GreedySynth$ to find a best lens in that class if
one exists ($\LC$),
using the examples $exs$ to filter out potential lenses that do not satisfy
the specification,
\item replaces $best$ with $\LC$, if $\LC$ has lower cost according
to our information-theoretic metric, and
\item adds the SREs derived from rewriting $S$ and $T$ ($\Expand(S,T)$)
  to the priority queue.
\end{itemize}
When the loop terminates, the search returns the globally best lens found
($best$).  Each subroutine of this algorithm will be explained in
further depth in the following subsections.

@@

The procedure \PCF{\Continue} returns false based on the how long the search has
been going, and how hard it expects the next class of problems to be. In
particular, if $\PCF{PQ.Peek}(pq) = (\Regex,\RegexAlt)$, that RE pair is at
distance $d$, the number of pairs in $pq$ at distance $d$ is $n$, and the
current best lens has cost $c$, then $\Call{Continue}{pq}$ checks whether $c > d
+ \Log_2(n)$. This termination condition is based around two primary principles:
do not search overly deep ($d$), and do not tackle a frontier that would take
too long to process ($\Log_2(n)$). The log of the frontier is included because
the frontier grows much faster than lens costs grow. Lens cost grows with the
log of the expected number of choices in a given lens (due to its information
theoretic basis), so the frontier calculation does too.

When $\Call{Continue}{pq}$ returns $\False$, the algorithm stops proposing
regular expression pairs, and instead returns to the user the best lens found
thus far. If the algorithm finds a bijective lens, which
has zero cost, it will immediately return.
%
\subsection{Stochastic DNF Regular Expressions}
\label{subsec:sdnfre}
The \GreedySynth subroutine converts input SREs into a temporary pseudo-normal
form, \emph{Stochastic DNF regular expressions} (\SDNFREabbrev{}s).
\SDNFREabbrev{}s normalize across many of the star-semiring equivalences -- if
two regular expressions are equal modulo differences in associativity,
commutativity, or distributivity, their corresponding SREs are syntactically
equal.

@@

The algorithm for converting a stochastic regular expressions into its
corresponding stochastic DNF regular expressions, written
$\ToDNFRegexOf{\Regex}$, is defined below. This conversion relies on operators
defined in Figure~\ref{fig:dnf-regex-functions}.
\[
  \begin{array}{rcl@{\hspace*{3em}}rcl}
    \ToDNFRegexOf{\String} 
    & = 
    & \DNFOf{(\SequenceOf{\String},1)}

@@

  \end{array}
\]
After this conversion has taken place, the sequences are ordered (thereby
normalizing for commutativity differences).
We have proven that this conversion respects languages and probability
distributions.

@@

@@

\paragraph*{\ToStochastic} With stochastic DNF regular expressions and
\ToDNFRegex defined, it is easier to explain \ToStochastic, the function that
converts regular expressions into stochastic regular expressions. If $\Regex$ is
a stochastic regular expression generated by \ToStochastic, then when put into
\SDNFREabbrev form,
$\ToDNFRegexOf{\Regex} = \DNFOf{(\Sequence_1,\frac{1}{n}) \DNFSep \ldots \DNFSep
  (\Sequence_n,\frac{1}{n})}$ for some sequences $\Sequence_1\ldots\Sequence_n$,
and every stochastic atom generated by \ToStochastic is
$\PRegexStar{\DNFRegex}{.8}$. In particular, $\ToDNFRegex$ generates regular
expressions whose DNF form gives equal probability to all sequence subcomponents
of the DNF regular expression, and gives iterations a .8 chance to continue
iterating. In our experience generating random strings from regular expressions,
these probabilities provide good distributions of strings---stars are iterated 4
times on average, and no individual choice in a series of disjunctions is chosen
disproportionately often.

Even though our generated probability distribution works well in most
situations, it is not perfect. Consider synthesizing a lens between the formats
shown in Figure~\ref{fig:minimized-representations}. While salary information is
present in \lstinline{emp_salaries}, but not in \lstinline{emp_insurance}, the
algorithm might spend a long time (fruitlessly) trying to construct lenses that
transform the salary to information present in \lstinline{emp_insurance}.

One solution would be to cut off the search early. However, then one runs into
the opposite problem: in other scenarios salary information may be present in
the other format, but it may well take quite a bit of work to find a
transformation that connects the salary in \lstinline{emp_salaries} to the
salary information in the second format. In such a scenario, early termination
may leave salary information discarded where it should be retained.

We can solve both problems by allowing users to augment the specifications with
\emph{relevance annotations}. Sometimes, users have external knowledge that certain
information appears exclusively in one format or the other or conversely appears
in both formats. By communicating this knowledge to the synthesis algorithm
through relevance annotations, users can force the synthesis of lenses that
discard or retain certain information.
The first annotation, $\SkipOf{\Regex}$, says the information of
$\Regex$ appears only in $\Regex$, and can safely be
projected. The second annotation, $\SRequireOf{\Regex}$, says the information of
$\Regex$ appears in the other format and cannot be discarded.

For example, the user can easily recognize that salary information of
\lstinline{emp_salaries} is not present in \lstinline{emp_insurance}. By
annotating the \lstinline{salary} field as \lstinline{skip(salary)}, users can
add this knowledge to the specification to optimize the search. In practice,
we define \Skip by setting the information content of $\Skip(\Regex)$ to zero.
\begin{gather*}
  \EntropyOf{\Skip(\Regex)} = 0
\end{gather*}

Similarly, users can recognize that employee names are present in both files. By
annotating instances of \lstinline{name} as \lstinline{require(name)}, users can
add this knowledge to the specification to optimize the search and enforce that
the returned lens that retains \lstinline{name} information. In practice, we
define \SRequire by setting the cost of any lens that loses ``required''
information to infinity.
%
\begin{center}
  \begin{tabular}{rcl}
    $\REntropyOf{\SRequireOf{\RegexAlt} \Given \Lens, \Regex}$
    & =
    & $\begin{cases}
      \infty & \text{if }\REntropyOf{\SRequireOf{\RegexAlt} \Given \Lens, \Regex}
      \neq 0\\
    0 & \text{otherwise}
  \end{cases}$\\

    $\REntropyOf{\RegexAlt \Given \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}, \Regex}$
    & =
    & $\begin{cases}
      \infty & \vspace*{-.3em}\text{if $\RegexAlt$ contains $\SRequireOf{\RegexAlt'}$ as a}\\
      & \text{subexpression, where $\EntropyOf{\RegexAlt'} \neq 0$}\\
    [\EntropyOf{\RegexAlt},\EntropyOf{\RegexAlt}] & \text{otherwise}
  \end{cases}$\\
  \end{tabular}
\end{center}

@@

The structure of symmetric DNF lenses exactly mirrors that of SDNF REs. A
symmetric DNF lens $sdl$ is a union of symmetric sequence lenses, $ssql_1 \ldots
ssql_p$, a symmetric DNF sequence lens is a concatenation of symmetric atom
lenses, $sal_1 \ldots sal_n$, and a symmetric atom lens $sal$ is an iteration of
a symmetric DNF lens.

Just as we analyzed the information content of ordinary regular expressions, we
can analyze the information content of DNF regular expressions. As before, we
use $\REntropyOf{\DNFRegexAlt \Given \SDNFLens, \DNFRegex}$ to calculate bounds
on the expected amount of information required to recover a string in
$\DNFRegexAlt$ from a string in $\DNFRegex$, synchronized by $\SDNFLens$. We use
the function $\LEntropyOf{\DNFRegex \Given \SDNFLens, \DNFRegexAlt}$ to
calculate bounds on the expected amount of information required to recover a
string in $\DNFRegex$ from a string in $\DNFRegexAlt$, synchronized by
$\SDNFLens$.

@@

\paragraph*{Compositional Synthesis} Compositional synthesis allows us to use
previously defined lenses, whether they are previously defined by users or by
synthesis. As the synthesizer processes a Boomerang file, it accumulates
lens definitions and their types.  It tackles synthesis problems one after another
and there may be many such problems in a given program.
During synthesis, if an existing lens has the right type and agrees with the examples.
\GreedySynth will use it.  Consequently, if a large synthesis problem cannot be solved all
at once, a user can first generate subproblems for parts of a data format, and the
larger problem can subsequently use solutions to those subproblems.  In our experience,
this is a very powerful tool that allows synthesis to tackle problems of arbitrary
complexity.
% end synthesis
